---
title: "Untitled"
output: 
  revealjs::revealjs_presentation:
    theme: sky
    highlight: pygments
    center: true
    transition: fade
    self_contained: false
    reveal_options:
      slideNumber: true
      previewLinks: true
---



# 1st: Set Up

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.






## Display modes

* 'f': enable fullscreen mode.
* 'o': enable overview mode.
* Pressing Esc exits all of these modes






## Appearance and style {data-transition="slide-in fade-out"}

theme specifies the theme to use for the presentation (available themes are "default", "simple", "sky", "beige", "serif", "solarized", "blood", "moon", "night", "black", "league", and "white").

highlight specifies the syntax highlighting style. Supported styles include "default", "tango", "pygments", "kate", "monochrome", "espresso", "zenburn", and "haddock". Pass null to prevent syntax highlighting.

center specifies whether you want to vertically center content on slides (this defaults to false).

smart 指示是否产生印刷正确的输出，将直引号转换为弯引号，---转换为破折号，-转换为破折号，以及...转换为椭圆。 请注意，默认情况下启用智能。






## 过渡样式 transition style {data-transition="zoom"}


transition specifies the visual effect when moving between slides. Available transitions are "default", "fade", "slide", "convex", "concave", "zoom" or "none". 过渡指定在幻灯片之间移动时的视觉效果。 可用的过渡是“默认”，“淡入淡出”，“滑动”，“凸出”，“凹入”，“缩放”或“无”。

background_transition specifies the background transition effect when moving between full page slides. Available transitions are "default", "fade", "slide", "convex", "concave", "zoom" or "none". 指定在整页幻灯片之间移动时的背景过渡效果





## Slide backgrounds

* ## CSS color background {data-background=#ff0000}
* ## Full size image background {data-background="background.jpeg"}
* ## Video background {data-background-video="background.mp4"}
* ## A background page {data-background-iframe="https://example.com"}




## Slide with Bullets {data-transition-speed="fast"}

- Bullet 1
- Bullet 2
- Bullet 3







# In the evening

## Slide with R Code and Output

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```

