---
title: "Components"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    navbar:
      - { title: "About", href: "https://example.com/about" }   ## navbar选项将自定义链接添加到导航栏
    social: [ "twitter", "facebook", "menu" ]                   ## social option can include any number of the following services: "facebook", "twitter", "google-plus", "linkedin", and "pinterest". You can also specify "menu" to provide a generic sharing drop-down menu that includes all of the services.
---

```{r setup, include=FALSE}
library(flexdashboard)
library(dygraphs)


# these computing functions are only toy examples
# valueBox（）函数以发出一个值并指定一个图标
computeArticles = function(...) return(45)
computeComments = function(...) return(126)
computeSpam = function(...) return(15)


## Gauges 显示指定范围内仪表上的值
## 三个必需的参数：value，min和max
## specify a set of three value ranges (success, warning, and danger)
```





Value Boxes {data-icon="fa-signal"}
===================================== 

### Articles per Day

```{r}
articles = computeArticles()
valueBox(articles, icon = "fa-pencil")
```

### Comments per Day

```{r}
comments = computeComments()
valueBox(comments, icon = "fa-comments")
```

### Spam per Day

```{r}
spam = computeSpam()
valueBox(
  spam, icon = "fa-trash",
  color = ifelse(spam > 10, "warning", "primary")
)
```











Gauges {data-icon="fa-signal"}
===================================== 

### Contact Rate

```{r}
gauge(91, min = 0, max = 100, symbol = '%', gaugeSectors(
  success = c(80, 100), warning = c(40, 79), danger = c(0, 39)
))
```

### Average Rating 

```{r}
gauge(37.4, min = 0, max = 50, gaugeSectors(
  success = c(41, 50), warning = c(21, 40), danger = c(0, 20)
))
```

### Cancellations

```{r}
gauge(7, min = 0, max = 10, gaugeSectors(
  success = c(0, 2), warning = c(3, 6), danger = c(7, 10)
))
```








   
Text annotations {data-icon="fa-table"}
=====================================     

Monthly deaths from bronchitis, emphysema and asthma in the
UK, 1974–1979 (Source: P. J. Diggle, 1990, Time Series: A
Biostatistical Introduction. Oxford, table A.3)


Row {data-height=600}
-------------------------------------

### All Lung Deaths

```{r}
dygraph(ldeaths)
```

Row {data-height=400}
-------------------------------------

### Male Deaths

```{r}
dygraph(mdeaths)
```

> Monthly deaths from lung disease in the UK, 1974–1979

### About dygraphs

This example makes use of the dygraphs R package. The dygraphs
package provides rich facilities for charting time-series data 
in R. You can use dygraphs at the R console, within R Markdown
documents, and within Shiny applications.

